---
title: "lucchese_a4"
author: "veronica m lucchese"
date: "2023-10-25"
output: html_document
---

```{r}
library(tidyverse)
library(DBI)
library(dbplyr)
library(bigrquery)

#After you have initialized a project, paste your project ID into the following chunk.
project <- "sonic-mile-403119"


#connect to a public database, the Chicago crime database, which has data on crime in Chicago
con <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "chicago_crime",
  billing = project
)

#look at the info about in this database using dbListTables.
dbGetInfo(con)

#look at the available tables in this database using dbListTables
dbListTables(con)

```
```{r}
# Seeing what columns are available in the crime data table
columns <- dbGetQuery(con, "
  SELECT *
  FROM `bigquery-public-data.chicago_crime.crime`
  LIMIT 1
")

print(columns)

```

```{sql connection= con, message = FALSE}
#Write a first query that counts the number of rows of the `crime` table in the year 2016.

SELECT COUNT(*) as row_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE year = 2016;
```


```{sql connection= con, message = FALSE}
#count the number of arrests grouped by primary_type in 2016. Note that is a somewhat similar task as above, with some adjustments on which rows should be considered. Sort the results, i.e. list the number of arrests in a descending order.
SELECT primary_type, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE year = 2016
  AND arrest = TRUE
GROUP BY primary_type
ORDER BY arrest_count DESC;
```

#Count the number of arrests grouped by hour of the day in 2016. You can extract the latter information from date via EXTRACT(HOUR FROM date). Which time of the day is associated with the most arrests?

# 10:00 o'clock is associated with the most arrests.
```{sql connection= con, message = FALSE}

SELECT EXTRACT(HOUR FROM date) AS hour_of_day, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE year = 2016 AND arrest = TRUE
GROUP BY hour_of_day
ORDER BY arrest_count DESC;

```

#Focus only on HOMICIDE and count the number of arrests for this incident type, grouped by year. List the results in descending order.
```{sql connection= con, message = FALSE}

SELECT EXTRACT(YEAR FROM date) AS year, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE primary_type = 'HOMICIDE'
  AND arrest = TRUE
GROUP BY year
ORDER BY arrest_count DESC;

```

#Find out which districts have the highest numbers of arrests in 2015 and 2016. That is, count the number of arrests in 2015 and 2016, grouped by year and district. List the results in descending order.

```{sql connection= con, message = FALSE}

SELECT EXTRACT(YEAR FROM date) AS year, district, COUNT(*) AS arrest_count
FROM `bigquery-public-data.chicago_crime.crime`
WHERE EXTRACT(YEAR FROM date) IN (2015, 2016) AND arrest = TRUE
GROUP BY year, district
ORDER BY arrest_count DESC;

```

```{r}
###writing queries from within R via the DBI package
library(DBI)

#Create a query object that counts the number of arrests grouped by primary_type of district 11 in year 2016. The results should be displayed in descending order.

query <- "
  SELECT primary_type, COUNT(*) AS arrest_count
  FROM `bigquery-public-data.chicago_crime.crime`
  WHERE district = 11 AND EXTRACT(YEAR FROM date) = 2016 AND arrest = TRUE
  GROUP BY primary_type
  ORDER BY arrest_count DESC"

# Execute the query
result <- dbGetQuery(con, query)

# Display the results
print(result)
```

```{r}
library(DBI)
library(dbplyr)
library(bigrquery)

#write code that counts the number of arrests grouped by primary_type of district 11 in year 2016, now using the dbplyr package. For this, you need to first map the crime table to a tibble object in R.

# I had issues with the DBPLYR package. Below is the commented code I would have used to do this prompt.
# Map the chicago_crime table to a tibble
#crime_db <- tbl(con, "crime")

# Filter data for district 11 and year 2016
#filtered_data <- crime_db %>%
#  filter(district == 11, year == 2016)

# Group by primary_type and count the number of arrests
#arrest_count <- filtered_data %>%
#  group_by(primary_type) %>%
#  summarise(arrests = n())

# Print the result
#print(arrest_count)


# SQL query to count the number of arrests grouped by primary_type for district 11 in the year 2016
query2 <- "
  SELECT primary_type, COUNT(*) as arrest_count
  FROM crime
  WHERE district = 11 AND year = 2016
  GROUP BY primary_type
  ORDER BY arrest_count DESC
"

# Execute the query
result2 <- dbGetQuery(con, query2)

print(result2)
```
```{r}
#Again, count the number of arrests grouped by primary_type of district 11 in year 2016, now using dplyr syntax.

# I had issues with the DBPLYR package. Below is the commented code I would have used to do this prompt.
# Map the chicago_crime table to a tibble
#crime_db <- tbl(con, "crime")

# Filter data for district 11 and year 2016
#filtered_data <- crime_db %>%
#  filter(district == 11, year == 2016)

# Group by primary_type and count the number of arrests
#arrest_count <- filtered_data %>%
#  group_by(primary_type) %>%
#  summarise(arrests = n())

# Print the result
#print(arrest_count)

query2 <- "
  SELECT primary_type, COUNT(*) as arrest_count
  FROM crime
  WHERE district = 11 AND year = 2016
  GROUP BY primary_type
  ORDER BY arrest_count DESC
"

# Execute the query
result2 <- dbGetQuery(con, query2)

print(result2)

```

```{r}
#Count the number of arrests grouped by primary_type and year, still only for district 11. Arrange the result by year.

# Filter data for district 11
#filtered_data <- crime_db %>%
#  filter(district == 11)

# Group by primary_type and year, and count the number of arrests
#arrest_count <- filtered_data %>%
 # group_by(primary_type, year) %>%
 # summarise(arrests = n())

# Arrange the result by year
#arrest_count <- arrest_count %>%
#  arrange(year)

# Print the result
#print(arrest_count)

query3 <- "
  SELECT primary_type, EXTRACT(YEAR FROM date) AS year, COUNT(*) as arrest_count
  FROM crime
  WHERE district = 11 
  GROUP BY primary_type, year
  ORDER BY arrest_count DESC
"

#Assign the results of the query above to a local R object.
# Execute the query
result3 <- dbGetQuery(con, query3)


#Confirm that you pulled the data to the local environment by displaying the first ten rows of the saved data set.
print(result3)


```

```{r}
#Close the connection.
dbDisconnect(con)
```